---
title: "R Packages"
subtitle: "How do I convert a script to a package?"
author: "Jon Harmon"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    includes:
      after_body: insert-logo.html
---

```{r libraries, include = FALSE}
library(xaringanthemer)
```

```{r xaringan-themer, include = FALSE}
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

# Vignette-Driven Development: What?

- **Vignette** = "long-form guide to your package."
- Rmarkdown doc.
- Shows how to solve a problem using your package.
- I do my day-to-day work as vignettes.

---

# Vignette-Driven Development: How?

- *Start* each analysis as a vignette (new or existing package).
- **Separate** analysis from helper functions.
- **Move** helper functions into package.
- **Reuse** helper functions.

---

# Let's Try!

- I'll build a package by converting the script in Chapter 6 into a vignette!

---

# Create the Package

- `usethis::create_package("../within")`
- Build > More > Configure Build Tools...
- Clean the DESCRIPTION (will discuss more in a few weeks)

```
Package: within
Title: An Example of Vignette-Driven Package Development
Version: 0.0.0.9000
Authors@R: person(given = "Jon",
    family = "Harmon",
    role = c("aut", "cre"),
    email = "jonthegeek@gmail.com",
    comment = c(ORCID = "0000-0003-4781-4346"))
Description: This package was built to demonstrate building a package starting 
    with a repetitive task in a vignette.
License: MIT + file LICENSE
Encoding: UTF-8
LazyData: true
Roxygen: list(markdown = TRUE)
RoxygenNote: 7.1.1
```

---

# Set Up GitHub

- `usethis::use_git()`
- `usethis::use_github()`

---

# Start the Vignette

- `usethis::use_vignette("swim-data", "Working with Swim Data")`
- Start explaining the process in the vignette:

```
We collect reports from people who went for a swim by asking this question:

> Where did you swim and how hot was it outside?

The data usually comes as a CSV file, which we read into a data frame.
```

- In bookclub project, copy swim.csv to within/vignettes dir.
- ctrl-shift-B to build package (often!).

---

# Code Blocks

- r shift-tab to insert an R code block.
- Copy the contents of bravo-baseline.R from bookclub proj to the vignette.
- Fully qualify all function calls.
- Remove `library` call.

---

# Test Often!

- ctrl-shift-K
- {sad trombone}
- `usethis::use_pipe()`
- ctrl-shift-B
- ctrl-shift-K

---

# Organize

- Split into clean-data and write-data blocks.
- Create file for the first block: `usethis::use_r("clean-data")`

---

# Functionize clean-data

- `dat <- dat %>% localize_beach()`
- Write `localize_beach` function in `cleaning-helpers.R` (ctrl-alt-x).
- ctrl-shift-B
- ctrl-shift-K
- {sad trombone}
- `#' @export` (we'll learn what this is next week)

---

# Functionize clean-data (continued)

- `%>% celsify_temp()`
- `celsify_temp` in `cleaning-helpers.R`
- ctrl-shift-B
- ctrl-shift-K

---

# Functionize write-data

- `usethis::use_r("write-data")`
- `outfile <- make_outfile_path(infile)`
- `make_outfile_path` function.
- `.pretty_now` and `.clean_path` functions.
 - They put the Sys.time outside to prove a point, but don't do that.
- ctrl-shift-B
- ctrl-shift-K

---

# A Second Vignette

- `usethis::use_vignette("thanking-submitters", "Thanking Submitters")`
- Wrap functions into `clean_swim_data`.

---

# Check Learning Objectives

- Understand how to **organize scripts** into **objects** and **functions.** `r emo::ji("heavy_check_mark")`
- Recognize how to **separate analysis from helper functions.** `r emo::ji("heavy_check_mark")`
- **Recognize errors** that can occur when converting a script to a package. `r emo::ji("heavy_check_mark")`
- Recognize that **top-level R code** is executed when a package is **built.**
- Be very mindful about **changing a user’s overall state** from within a package.
- Define a **vignette** as a **long-form guide** to your package. `r emo::ji("heavy_check_mark")`
- Create a vignette with `usethis::use_vignette()`. `r emo::ji("heavy_check_mark")`
- Recognize the **metadata fields** in a vignette. `r emo::ji("heavy_check_mark")`
- Remember that **markdown** allows you to format your vignette.
- Understand how to **intermingle R code** and text in vignettes. `r emo::ji("heavy_check_mark")`
- Remember how to **view your vignette** as it will appear to a user. `r emo::ji("heavy_check_mark")`

---

# Check Learning Objectives

- Understand how to **organize scripts** into **objects** and **functions.** `r emo::ji("heavy_check_mark")`
- Recognize how to **separate analysis from helper functions.** `r emo::ji("heavy_check_mark")`
- **Recognize errors** that can occur when converting a script to a package. `r emo::ji("heavy_check_mark")`
- Recognize that **top-level R code** is executed when a package is **built.** `r emo::ji("heavy_check_mark")`
- Be very mindful about **changing a user’s overall state** from within a package. `r emo::ji("heavy_check_mark")`
- Define a **vignette** as a **long-form guide** to your package. `r emo::ji("heavy_check_mark")`
- Create a vignette with `usethis::use_vignette()`. `r emo::ji("heavy_check_mark")`
- Recognize the **metadata fields** in a vignette. `r emo::ji("heavy_check_mark")`
- Remember that **markdown** allows you to format your vignette. `r emo::ji("heavy_check_mark")`
- Understand how to **intermingle R code** and text in vignettes. `r emo::ji("heavy_check_mark")`
- Remember how to **view your vignette** as it will appear to a user. `r emo::ji("heavy_check_mark")`
