---
title: "R Packages"
subtitle: "How do I include things like data and do obscure package-y things?"
author: "Jon Harmon"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    includes:
      after_body: insert-logo.html
---

```{r libraries, include = FALSE}
library(xaringanthemer)
```

```{r xaringan-themer, include = FALSE}
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

# Chapter 14 learning objectives

* List the three main ways to include data in a package.
 * Exported data: `data/`
 * Internal data: `R/sysdata.rda`
 * Raw data (for export): `inst/extdata`
* Define `LazyData`.
 * Data isn't loaded until used.
* Use `usethis::use_data_raw()` to document how a dataset was created.
* Use `usethis::use_data()` to add data to a package.
* Use roxygen2 to document a dataset in a package.
* Use `@format` to  provide a data dictionary for a dataset.
* Use `@source` to provide details of where you got the data.
* Use `usethis::use_data(..., internal = TRUE)` to add internal data to a package.
* Provide raw data to a package user in the `inst/extdata` folder.
* Load raw data with `system.file`.
* Note where data for tests should be saved.
* Note the options for data in vignettes.
* Use `tools::checkRdaFiles()` and `tools::resaveRdaFiles()` to optimize compression for package data.

---

# Chapter 15 learning objectives

* Use `usethis::use_rcpp()` or `usethis::use_cpp11` to set up a package for C++.
* Understand the workflow for working with C++ in R.
* Understand how to document `Rcpp` code.
* Understand how to export C++ code so that it can be called by other C++ code.
* Understand how to import C++ code from another package.
* Note the best practices for C++ code in packages.
* Remember that this book has more information about working with C code.
* Recognize workflows for debugging compiled code.
* Remember that this book has more information about working with Makefiles.
* Recall that compiled code must work on Linux, Mac, and Windows, so it must be checked carefully before submitting to CRAN. 

We also mentioned [Efficient R Programming](https://csgillespie.github.io/efficientR/index.html)

---

# Chapter 16-17 learning objectives

* Use `system.file()` to load a file included in `inst`.
* Use `citation()` to find citation information for R and R packages.
* Use `usethis::use_citation()` to customize the citation for your package.
* Recognize how to include scripts from other programming languages.
* Recognize that there are other optional components in packages, stored in `demo`, `exec`, `po`, and `tools`.
 * The [lubridate](https://github.com/tidyverse/lubridate) package has translations in `po`!
 * `po` stands for "portable object".
