---
title: "R Packages"
subtitle: "Testing and Automated Checking"
author: "David Edison"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    includes:
      after_body: insert-logo.html
---

```{r libraries, include = FALSE}
library(xaringanthemer)
```

```{r xaringan-themer, include = FALSE}
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

# Chapter 12: Testing

## Why should we write tests?

- Fewer bugs
- Better code structure
- Easier restarts
- Robust code

---

# Chapter 12: Testing

## Workflow

- testthat package (version 3.0.0 out on CRAN!)
- `usethis::use_testthat` function sets up most of the structure for you
- Book recommends testthat (for obvious reasons), but other testing frameworks exist (e.g. tinytest)

---

# Chapter 12: Testing

## Structure

- A test file lives in the `tests/testthat` directory
- Create a new test with `usethis::use_test` function. It is recommended to use the same name as the corresponding file in the `R` directory (use descriptive filenames -- despite what the book says, `context` is deprecated)
- An *expectation* is the atom of testing
- A *test* groups together multiple expectations. Also referred to as a *unit test* since they test one "unit" of functionality
- A *file* groups together multiple related tests

---

# Chapter 12: Testing

## Expectations

- All expectation functions in testthat have 2 arguments: the actual result, and the expectation
- If they don't agree, testhat throws an error
- New to testthat 3.0.0, snapshot tests can be used to visually inspect output
- Examples:
```
expect_equal
expect_error
expect_s3_class
expect_match
expect_lt
expect_snapshot_output
```
---

# Chapter 12: Testing

## Writing Tests

- Each test should have an informative name and cover a single unit of functionality
- Focus on testing the external interface to your functions
- Test each behavior in only one test
- Avoid testing simple code you're confident will work
- Always write a test when you discover a bug
- You can use `skip` function to conditionally skip a test
- Sometime it's useful to build your own testing tools to reduce duplication

---

# Chapter 12: Testing

## CRAN Notes

- Tests should run in under a minute (you can add a call to `skip_on_cran` before long running tests)
- Tests are always run in the English Language
- Be careful about testing things that are likely to be variable on CRAN machines (e.g. testing how long something takes, numerical precision)

---

# Chapter 19: Automated Checking 

## Worflow

- Automated checking automatically checks code for common problems
- `R CMD check` is termal command or use `devtools::check` (Ctrl/Cmd + Shift + E in RStudio)
- `R CMD check` returns 3 types of messages
  - `ERRORS` (severe problems, definitely fix)
  - `WARNINGS` (likely problems you must fix if submitted to CRAN)
  - `NOTES` (mild problems, should strive to eliminate in submitting to CRAN)
  
---

# Chapter 19: Automated Checking 

## Checks

- Metadata 
- Package Structure 
- `DESCRIPTION` file meta-information
- Package dependencies
- CRAN feasibility
- Namespace
- R code
- Data
- Documentation
- Demos
- Compiled code
- Tests
- Vignettes

---

# Chapter 19: Automated Checking 

## Continuous Integration

- Automatically run `devtools::check` every time code is pushed to github
- Basic use
  - `usethis::use_travis`, `usethis::use_circleci`, `usethis::use_appveyor`
  - Set up a config file
  - Commit and push to github
  - Wait a few minutes for results to be emailed
- You can use CI tooling to do other cool stuff like replublish a book every time source changes or building vignettes and publishing them to a website

---

class: middle, center

# [Demo](https://github.com/d-edison/testdemo)
