---
title: "R Packages"
subtitle: "RPkgs - How Should I Work On My Package?"
author: "Kevin Gilds"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    includes:
      after_body: insert-logo.html
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      comment = "#>")

library(xaringanthemer)

```


```{r xaringan-themer, include = FALSE}
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```



## Outline

ch 5 and ch 18, cover powerful work tools in 

- RStudio Projects & Devtools
- Git & GitHub


---

## Chapter 5

- Naming Conventions & Practices `available::available()`
- Package Creation with `usethis:: create_package()`
    
    
    - R directory

    - DESCRIPTION file
    
    - NAMESPACE file

- Rstudio projects
  - Each project is isolated, with a file browser and working directory pre set
  - Keyboard shortcuts
- working directory and file path discipline

  - Don't fiddle with `setwd()` in the scripts in your package
  
  - write paths that convey intent explicitly
  
    - Path helpers: `testthat::test_path()`
    `fs::path_package()`
---

## Chapter 5 continued

  - `devtools::load_all()` simulates the package environment-you can check if your code works
    - quickly iterate
    - call functions from other packages that you have imported
      - Keyboard shortcut: Cmd+Shift+L (macOS), Ctrl+Shift+L (Windows, Linux)
      - Build pane more menu


---

_Practical advice_

- pick a name easy to Google

- pick a name that is easy to pronounce



```{r eval = FALSE}
library(available)

available::available("Rduck")


```

---

# Bare Bones R package


```{r eval=FALSE}
usethis::create_package("Rduck")
```


- R directory

- DESCRIPTION file

- NAMESPACE file


---


# Fundamental development workflows


**Learning objectives:**

* Recall the rules for R package names.
* Use `available::available()` to check whether a package name works.
* Use `usethis::create_package()` to create a package and the associated RStudio Project.
* List key benefits of working with RStudio Projects.
* Recognize that exercising path discipline helps you avoid errors as your package changes states.
* List the benefits of `devtools::load_all()`.
* Remember how to quickly call `devtools::load_all()` in RStudio.


---

# Git and GitHub

**Learning objectives:**

* Understand that Git is a version control system.
* Understand the difference between a local repository and a remote repository.
* Understand how to commit your work in Git.
* Understand how and when to push changes from one repository to another.
* Understand how and when to pull changes from one repository to another.
* Understand how to request that someone else pull your changes into their repository.
* Know some places to look for more information.

---

---
## Git

- Experience is key to using and valuing git
    - Git has saved me countless hours
    - Revert has saved me as well. 
    

- Git tracks your changes 
    - Helpful in debugging and gives you a history of the project
    

- Git branches
  - Lets you play and see if 'something will work' without losing what does work. 
    

---
  - The five components of a commit
    1. SHA unique identifier
    2. changeset--the files that were modified
    3. commit message--human readable
    4. parent-the prior commit
      - none for initial commits
      - merged have two parents
    5. the author

- Two stages of a commit
  - identify what files will be associated with the commit
  - commit the the staged files


---
# Best Practices for Git commits

The art of the commit.  

- Each commit should be minimal but complete
    - minimal: focus on a single problem
    - complete if you fix a bug include a unit test with the commit! 


- Commit messages
  - be concise but meaningful
  - focus on the why!  
      - 6 months from now you will see what you did but you may not know why you made that change. 

---
## Resources

Git Resources: 

- Jenny Bryan's [Happy Git With R](https://www.happygitwithr.com) (comprehensive use of Git with R)
  - Spend time here to get set up with github (Keys & tokens) This makes using usethis::use_github() so easy!
  
- [OhShitGit Zine](https://jvns.ca/blog/2018/10/27/new-zine--oh-shit--git-/) - Julia Evans / Katie Sylor-Miller -> I find Julia Evans' zines so incredibly useful - also check out her zines on SQL and (soon) CSS 

- [OhShitGit.com](https://ohshitgit.com) by Katie Sylor-Miller (alternatively, DangitGit.com for non-swearing)

- Git Explorer [https://gitexplorer.com/](https://gitexplorer.com/)

- Pro Git Book [Git Book](https://git-scm.com/book/en/v2) by Scott Chacon and Ben Straub


- Git Remotes [Git Remotes](https://youtu.be/Gg4bLk8cGNo) This is a nice video tutorial on Forking and Pull Requests by David Mahler


---








